---
title: "Script 2 – Model vs. mooring near-bottom currents (Figures 5, S2, S3)"
author: "Almog Haim – MEDsLab, Bar-Ilan University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

# Overview

This script compares the near-bottom currents of the hydrodynamic model
(Dec 2011 – Nov 2012, hourly) with the mooring observations, starting from
the **raw files**, and reproduces:

| Figure | Content |
|---|---|
| 5 | Mean current speed and persistence: model vs. mooring |
| S2 | Principal axes of the current: model vs. mooring |
| S3 | Full-year modelled near-bottom currents (stick plots) |

The model is compared with each mooring over the same calendar period
(the mooring dates transferred to the model year 2012), after both were
averaged to hourly values.

Code consolidated from the original model–mooring analysis script.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)   # dplyr, tidyr, ggplot2, readr, stringr, purrr
library(lubridate)   # date-time parsing
library(patchwork)   # multi-panel figure assembly
```

# 1. Input files and settings

All paths are relative to this script (the `scripts` folder). The mooring
files are the same raw current-meter files used in Script 1 (`data/moorings`).
The hydrodynamic-model output (hourly near-bottom currents at the 18 model
stations, December 2011 – November 2012) is in `data/model`, stored as a
gzip-compressed CSV that `read_csv()` reads directly.

```{r paths}
mooring_dir <- file.path("..", "data", "moorings")
model_file  <- file.path("..", "data", "model",
                         "near_bottom_currents_model_2011_2012.csv.gz")

mooring_files <- tribble(
  ~Canyon,    ~Location, ~Depth_m, ~file,
  "Eilat",    "In",  100, file.path(mooring_dir, "eilat_100m_in_current.csv"),
  "Eilat",    "Out", 100, file.path(mooring_dir, "eilat_100m_out_current.csv"),
  "Princess", "In",  100, file.path(mooring_dir, "princess_100m_in_current.csv"),
  "Princess", "Out", 100, file.path(mooring_dir, "princess_100m_out_current.csv"),
  "Shlomo",   "In",  100, file.path(mooring_dir, "shlomo_100m_in_current.csv"),
  "Shlomo",   "Out", 100, file.path(mooring_dir, "shlomo_100m_out_current.csv"),
  "Shlomo",   "In",  200, file.path(mooring_dir, "shlomo_200m_in_current.csv"),
  "Shlomo",   "Out", 200, file.path(mooring_dir, "shlomo_200m_out_current.csv")
)

# Output folder for figures, tables and the saved workspace
out_dir <- file.path("..", "outputs")
fig_dir <- out_dir
dir.create(fig_dir, showWarnings = FALSE)

missing <- c(mooring_files$file, model_file)[!file.exists(c(mooring_files$file, model_file))]
if (length(missing) > 0) stop("Missing input file(s):\n", paste(missing, collapse = "\n"))
```

```{r settings}
# Hours removed at the start and at the end of every mooring record
# (deployment and recovery of the moorings) – same rule as in Script 1
trim_hours <- 3

# Stick plots of the model year (Figure S3)
stick_bin_hours <- 24     # velocities averaged over daily bins
x_scale_factor  <- 4000   # horizontal stretch of the arrows (x in seconds, u in cm/s)

canyon_levels <- c("Eilat", "Shlomo", "Princess")
canyon_cols   <- c("Eilat" = "steelblue", "Shlomo" = "darkorange", "Princess" = "darkgreen")
source_cols   <- c("Mooring" = "#F8766D", "Model" = "#00BFC4")
season_cols   <- c("Winter" = "#D8C2C2", "Spring" = "#C9D7B8",
                   "Summer" = "#B8D4D8", "Autumn" = "#D2C3D8")

# ---------------------------------------------------------------------------
# Figure style – used by every figure so that all figures look the same
# ---------------------------------------------------------------------------
# Font: Times New Roman, requested through R's generic "serif" family.
# "serif" is understood by every R graphics device (screen, png, pdf,
# cairo_pdf, ragg) and is drawn as Times New Roman on Windows (Times on
# macOS). Asking for "Times New Roman" by name makes the standard pdf() and
# png() devices fail silently, which produces empty files.
font_family <- "serif"

# Text sizes (pt) shared by all figures
fig_sizes <- list(
  axis_title   = 18,   # axis titles
  axis_text    = 14,   # axis numbers
  strip        = 16,   # panel title strips
  legend_title = 16,
  legend_text  = 14,
  legend_key   = 1.3,  # size of the legend colour boxes (lines)
  tag          = 18,   # panel letters
  annotation   = 4.5   # text inside panels (mm, ggplot units)
)

theme_fig <- theme(
  text         = element_text(family = font_family),
  axis.title   = element_text(size = fig_sizes$axis_title),
  axis.text    = element_text(size = fig_sizes$axis_text, colour = "black"),
  strip.text   = element_text(size = fig_sizes$strip),
  legend.title = element_text(size = fig_sizes$legend_title, face = "bold"),
  legend.text  = element_text(size = fig_sizes$legend_text),
  plot.tag     = element_text(size = fig_sizes$tag, face = "bold")
)

# Text drawn inside the panels (geom_text / geom_label) uses the same font
update_geom_defaults("text",  list(family = font_family))
update_geom_defaults("label", list(family = font_family))

# Single y-axis title (left) and x-axis title (bottom) for a whole figure
add_figure_titles <- function(panels, y = NULL, x = NULL, x_hjust = 0.5,
                              size = fig_sizes$axis_title) {
  wrap_elements(panel = panels) +
    labs(tag = y, caption = x) +
    theme(text              = element_text(family = font_family),
          plot.tag          = element_text(angle = 90, size = size),
          plot.tag.position = "left",
          plot.caption      = element_text(size = size, hjust = x_hjust))
}

# Hide axis numbers but keep their space, so that panels stay evenly spaced
hide_axis_text <- function(axis = c("x", "y")) {
  axis <- match.arg(axis)
  if (axis == "x") {
    theme(axis.text.x = element_text(colour = "transparent"),
          axis.ticks.x = element_line(colour = "transparent"))
  } else {
    theme(axis.text.y = element_text(colour = "transparent"),
          axis.ticks.y = element_line(colour = "transparent"))
  }
}

# Save a figure as PDF (vector, embedded font), 300-dpi PNG and 300-dpi TIFF.
# Always pass the figure object explicitly: ggsave() without `plot =` saves
# ggplot2's "last plot", which for a patchwork figure can be one of its
# internal pieces (e.g. the empty legend panel).
save_figure <- function(plot, name, width, height, dpi = 300) {
  has_ragg   <- requireNamespace("ragg", quietly = TRUE)
  pdf_device <- if (capabilities("cairo")) cairo_pdf else "pdf"
  files <- file.path(fig_dir, paste0(name, c(".pdf", ".png", ".tiff")))
  start <- Sys.time() - 1

  ggsave(files[1], plot, width = width, height = height, device = pdf_device)
  ggsave(files[2], plot, width = width, height = height, dpi = dpi, bg = "white",
         device = if (has_ragg) ragg::agg_png else "png")
  ggsave(files[3], plot, width = width, height = height, dpi = dpi, bg = "white",
         device = if (has_ragg) ragg::agg_tiff else "tiff", compression = "lzw")

  # Windows cannot overwrite a file that is open in another program (image
  # viewer, PDF reader, Explorer preview pane). Check that every file was
  # really rewritten and say so if not.
  not_updated <- files[is.na(file.mtime(files)) | file.mtime(files) < start]
  if (length(not_updated) > 0) {
    warning("Not updated (probably open in another program - close it and run again):\n  ",
            paste(not_updated, collapse = "\n  "), call. = FALSE)
  } else {
    message("Saved: ", paste(basename(files), collapse = ", "))
  }

  set_last_plot(plot)   # so that a later ggsave() without `plot =` also saves this figure
  invisible(plot)
}


# Panel letters in the top-left corner of each panel
tag_inside <- function(x = 0.01, y = 0.99, size = fig_sizes$tag) {
  theme(plot.tag = element_text(size = size, face = "bold"), plot.tag.position = c(x, y))
}

# Axis labels with at most two decimals, trailing zeros dropped
label_2dec  <- function(x) sub("\\.?0+$", "", sprintf("%.2f", x))

# Legend drawn from a separate, otherwise empty plot, so that it appears once
# and looks the same in every ggplot2 version
# `scale` enlarges text and keys (e.g. 1.25 = 25 % larger than the default)
legend_panel <- function(p, scale = 1) {
  wrap_elements(full = p + theme_void() +
    theme(legend.position = "left",
          legend.key.size = unit(fig_sizes$legend_key * scale, "lines"),
          text         = element_text(family = font_family),
          legend.title = element_text(face = "bold", size = fig_sizes$legend_title * scale),
          legend.text  = element_text(size = fig_sizes$legend_text * scale)))
}
```

# 2. Helper functions

```{r helpers}
# Remove the first and the last `n_hours` of a record (deployment/recovery)
trim_deployment_edges <- function(df, n_hours = trim_hours) {
  df %>% filter(
    time >= min(time, na.rm = TRUE) + hours(n_hours),
    time <= max(time, na.rm = TRUE) - hours(n_hours)
  )
}

# One raw mooring file -> time (UTC, as the model), u (east) and v (north), cm/s
read_mooring <- function(file, canyon, location, depth) {
  # time read as text: automatic parsing drops the time of day at midnight
  read_csv(file, show_col_types = FALSE,
           col_types = cols(`ISO 8601 Time` = col_character(), .default = col_guess())) %>%
    transmute(
      time = ymd_hms(`ISO 8601 Time`, tz = "UTC"),
      Canyon = canyon, Location = location, Depth_m = depth,
      u = `Velocity-E (cm/s)`,
      v = `Velocity-N (cm/s)`
    ) %>%
    filter(!is.na(time)) %>%
    arrange(time) %>%
    trim_deployment_edges()
}

# Summary statistics of the current at each station
calculate_current_stats <- function(data, dataset_name) {
  data %>%
    group_by(Canyon, Location, Depth_m) %>%
    summarise(
      N             = n(),
      Duration_days = as.numeric(difftime(max(time), min(time), units = "days")),
      Mean_speed    = mean(speed, na.rm = TRUE),
      Median_speed  = median(speed, na.rm = TRUE),
      SD_speed      = sd(speed, na.rm = TRUE),
      P90_speed     = as.numeric(quantile(speed, 0.90, na.rm = TRUE)),
      Mean_U        = mean(u, na.rm = TRUE),
      Mean_V        = mean(v, na.rm = TRUE),
      Vector_mean_speed = sqrt(Mean_U^2 + Mean_V^2),
      Persistence   = Vector_mean_speed / Mean_speed,   # 1 = always the same direction
      North_pct     = mean(v > 0, na.rm = TRUE) * 100,
      South_pct     = mean(v < 0, na.rm = TRUE) * 100,
      East_pct      = mean(u > 0, na.rm = TRUE) * 100,
      West_pct      = mean(u < 0, na.rm = TRUE) * 100,
      .groups = "drop"
    ) %>%
    mutate(
      # Direction of the mean current vector (0 = North, 90 = East)
      Mean_direction_deg = (atan2(Mean_U, Mean_V) * 180 / pi + 360) %% 360,
      Dataset = dataset_name
    )
}

# Principal axis of the current fluctuations: first eigenvector of the u'-v'
# covariance matrix, as a compass bearing between 0 and 180 degrees
calculate_principal_axis <- function(data, dataset_name) {
  data %>%
    filter(!is.na(u), !is.na(v)) %>%
    group_by(Canyon, Location, Depth_m) %>%
    group_modify(~ {
      eig <- eigen(cov(cbind(.x$u - mean(.x$u), .x$v - mean(.x$v))))
      major <- eig$vectors[, 1]
      tibble(
        Principal_axis_deg     = (atan2(major[1], major[2]) * 180 / pi) %% 180,
        Variance_explained_pct = 100 * eig$values[1] / sum(eig$values),
        Major_axis_SD          = sqrt(eig$values[1]),
        Minor_axis_SD          = sqrt(eig$values[2]),
        Minor_major_ratio      = sqrt(eig$values[2] / eig$values[1])
      )
    }) %>%
    ungroup() %>%
    mutate(Dataset = dataset_name)
}

# Smallest difference between two directions (0–180) and two axes (0–90)
angular_difference <- function(a, b) abs((a - b + 180) %% 360 - 180)
axial_difference   <- function(a, b) abs((a - b + 90) %% 180 - 90)

# Meteorological seasons (Dec–Feb winter, Mar–May spring, ...)
make_season_bands <- function(time_vec) {
  years <- seq(year(min(time_vec)) - 1, year(max(time_vec)))
  map_dfr(years, function(y) tibble(
    season = c("Winter", "Spring", "Summer", "Autumn"),
    xmin = as.POSIXct(paste0(c(y, y + 1, y + 1, y + 1), c("-12-01", "-03-01", "-06-01", "-09-01")), tz = "UTC"),
    xmax = as.POSIXct(paste0(y + 1, c("-03-01", "-06-01", "-09-01", "-12-01")), tz = "UTC")
  )) %>%
    filter(xmax >= min(time_vec), xmin <= max(time_vec)) %>%
    mutate(xmin = pmax(xmin, min(time_vec)), xmax = pmin(xmax, max(time_vec)),
           season = factor(season, levels = names(season_cols)))
}
```

# 3. Load the data

```{r load-moorings}
mooring <- pmap_dfr(mooring_files, function(Canyon, Location, Depth_m, file) {
  read_mooring(file, Canyon, Location, Depth_m)
})

# Hourly means, to match the model resolution
mooring_hourly <- mooring %>%
  mutate(time = floor_date(time, "hour")) %>%
  group_by(Canyon, Location, Depth_m, time) %>%
  summarise(u = mean(u, na.rm = TRUE), v = mean(v, na.rm = TRUE), .groups = "drop") %>%
  mutate(speed = sqrt(u^2 + v^2), Source = "Mooring")

# Period of each mooring record (after trimming)
mooring_windows <- mooring %>%
  group_by(Canyon, Location, Depth_m) %>%
  summarise(Mooring_start = min(time), Mooring_end = max(time), N = n(), .groups = "drop") %>%
  mutate(Model_start = update(Mooring_start, year = 2012),   # same calendar dates,
         Model_end   = update(Mooring_end,   year = 2012))   # in the model year

mooring_windows
```

```{r load-model}
model <- read_csv(model_file, show_col_types = FALSE) %>%
  transmute(
    time = dmy_hm(DateTime, tz = "UTC"),
    Canyon, Location,
    Depth_m = TargetDepth_m,
    BottomDepth_m,
    u = U_cm_s, v = V_cm_s,
    speed = sqrt(u^2 + v^2),
    Source = "Model"
  )

# Model grid points at the mooring stations, full year and same season
model_matching <- model %>%
  semi_join(distinct(mooring_hourly, Canyon, Location, Depth_m),
            by = c("Canyon", "Location", "Depth_m"))

model_seasonal <- model_matching %>%
  inner_join(select(mooring_windows, Canyon, Location, Depth_m, Model_start, Model_end),
             by = c("Canyon", "Location", "Depth_m")) %>%
  filter(time >= Model_start, time <= Model_end) %>%
  select(-Model_start, -Model_end)
```

# 4. Model–mooring comparison tables

```{r stats}
comparison_long <- bind_rows(
  calculate_current_stats(mooring_hourly, "Mooring"),
  calculate_current_stats(model_seasonal, "Model_seasonal"),
  calculate_current_stats(model_matching, "Model_full")
)

comparison_compact <- comparison_long %>%
  select(Canyon, Location, Depth_m, Dataset, Mean_speed, Mean_direction_deg, Persistence) %>%
  pivot_wider(names_from = Dataset,
              values_from = c(Mean_speed, Mean_direction_deg, Persistence)) %>%
  left_join(distinct(model_matching, Canyon, Location, Depth_m, BottomDepth_m),
            by = c("Canyon", "Location", "Depth_m")) %>%
  mutate(
    Canyon = factor(Canyon, levels = canyon_levels),
    Bathymetry_difference = abs(Depth_m - BottomDepth_m),
    Speed_difference_pct  = 100 * (Mean_speed_Model_seasonal - Mean_speed_Mooring) / Mean_speed_Mooring,
    Direction_difference  = angular_difference(Mean_direction_deg_Model_seasonal,
                                               Mean_direction_deg_Mooring),
    Station = factor(paste0(Location, " ", Depth_m, " m"),
                     levels = c("In 100 m", "Out 100 m", "In 200 m", "Out 200 m"))
  ) %>%
  arrange(Canyon, Depth_m, Location)

comparison_compact %>%
  select(Canyon, Station, BottomDepth_m, Mean_speed_Mooring, Mean_speed_Model_seasonal,
         Mean_speed_Model_full, Speed_difference_pct, Persistence_Mooring,
         Persistence_Model_seasonal, Direction_difference) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  knitr::kable()

principal_comparison <- bind_rows(
  calculate_principal_axis(mooring_hourly, "Mooring"),
  calculate_principal_axis(model_seasonal, "Model")
) %>%
  pivot_wider(names_from = Dataset,
              values_from = c(Principal_axis_deg, Variance_explained_pct,
                              Major_axis_SD, Minor_axis_SD, Minor_major_ratio)) %>%
  mutate(Axis_difference_deg = axial_difference(Principal_axis_deg_Mooring,
                                                Principal_axis_deg_Model),
         Canyon = factor(Canyon, levels = canyon_levels)) %>%
  arrange(Canyon, Depth_m, Location)

principal_comparison %>%
  select(Canyon, Location, Depth_m, Principal_axis_deg_Mooring, Principal_axis_deg_Model,
         Axis_difference_deg, Variance_explained_pct_Mooring, Variance_explained_pct_Model) %>%
  mutate(across(where(is.numeric), ~ round(.x, 1))) %>%
  knitr::kable()
```

```{r export-tables}
write_csv(comparison_long,      file.path(fig_dir, "Table_Model_Mooring_comparison_long.csv"))
write_csv(comparison_compact,   file.path(fig_dir, "Table_Model_Mooring_comparison_compact.csv"))
write_csv(principal_comparison, file.path(fig_dir, "Table_Principal_axes.csv"))
```

# 5. Figure 5 – mean speed and persistence

```{r fig5, fig.width=15, fig.height=7}
station_shapes <- c("In 100 m" = 15, "Out 100 m" = 16, "In 200 m" = 17, "Out 200 m" = 18)

theme_scatter <- theme_classic() + theme_fig

scatter_panel <- function(x, y, lim, breaks, xlab, ylab) {
  ggplot(comparison_compact, aes(x = .data[[x]], y = .data[[y]],
                                 colour = Canyon, shape = Station)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.8, colour = "grey40") +
    geom_point(size = 5) +
    scale_colour_manual(values = canyon_cols, guide = "none") +
    scale_shape_manual(values = station_shapes, guide = "none") +
    scale_x_continuous(limits = lim, breaks = breaks, labels = label_2dec,
                       expand = expansion(mult = c(0, 0.02))) +
    scale_y_continuous(limits = lim, breaks = breaks, labels = label_2dec,
                       expand = expansion(mult = c(0, 0.02))) +
    coord_equal() +
    labs(x = xlab, y = ylab) +
    theme_scatter
}

max_speed <- max(c(comparison_compact$Mean_speed_Mooring,
                   comparison_compact$Mean_speed_Model_seasonal)) * 1.08

p_speed <- scatter_panel("Mean_speed_Mooring", "Mean_speed_Model_seasonal",
                         lim = c(0, max_speed), breaks = seq(0, 12, by = 3),
                         xlab = expression(Mooring~mean~speed~(cm~s^{-1})),
                         ylab = expression(Model~mean~speed~(cm~s^{-1})))

p_persistence <- scatter_panel("Persistence_Mooring", "Persistence_Model_seasonal",
                               lim = c(0, 1), breaks = seq(0, 1, by = 0.2),
                               xlab = "Mooring persistence", ylab = "Model persistence")

legend_fig5 <- legend_panel(
  ggplot(comparison_compact, aes(x = 0, y = 0)) +
    geom_point(aes(colour = Canyon), alpha = 0, size = 5) +
    geom_point(aes(shape = Station), alpha = 0, size = 5) +
    scale_colour_manual(values = canyon_cols, name = "Canyon:",
                        guide = guide_legend(order = 1, override.aes = list(alpha = 1, shape = 16, size = 7))) +
    scale_shape_manual(values = station_shapes, name = "Station:",
                       guide = guide_legend(order = 2, override.aes = list(alpha = 1, size = 7))),
  scale = 1.3
)

figure5 <- p_speed + p_persistence + legend_fig5 +
  plot_layout(widths = c(1, 1, 0.55)) +
  plot_annotation(tag_levels = list(c("A", "B", ""))) &
  tag_inside(0.02, 0.99)

figure5

save_figure(figure5, "Figure 5", width = 15, height = 7)
```

# 6. Figure S2 – principal axes

Each panel shows the principal axis of the mooring and of the model at one
station. The length of the lines is fixed; the percentage of variance
explained by each axis is given above the circle and the difference between
the two axes (Δθ) below it.

```{r figS2-axes, fig.width=16, fig.height=9}
s1_order <- tribble(
  ~Canyon,    ~Location, ~Depth_m,
  "Eilat",    "In",  100, "Eilat",    "Out", 100,
  "Princess", "In",  100, "Princess", "Out", 100,
  "Shlomo",   "In",  100, "Shlomo",   "Out", 100,
  "Shlomo",   "In",  200, "Shlomo",   "Out", 200
)

circle <- tibble(a = seq(0, 2 * pi, length.out = 361), x = sin(a), y = cos(a))
cross  <- tibble(x = c(-1, 0), xend = c(1, 0), y = c(0, -1), yend = c(0, 1))
compass <- tibble(label = c("N", "E", "S", "W"),
                  x = c(0, 1.15, 0, -1.15), y = c(1.15, 0, -1.15, 0))

axis_panel <- function(cy, loc, dep) {
  pc <- principal_comparison %>% filter(Canyon == cy, Location == loc, Depth_m == dep)
  axes <- tibble(Source = factor(c("Mooring", "Model"), levels = names(source_cols)),
                 deg = c(pc$Principal_axis_deg_Mooring, pc$Principal_axis_deg_Model)) %>%
    mutate(xend = 0.85 * sin(deg * pi / 180), yend = 0.85 * cos(deg * pi / 180))

  ggplot() +
    geom_path(data = circle, aes(x, y), colour = "grey70", linewidth = 0.4) +
    geom_segment(data = cross, aes(x = x, y = y, xend = xend, yend = yend),
                 colour = "grey80", linewidth = 0.3) +
    geom_segment(data = axes, aes(x = -xend, y = -yend, xend = xend, yend = yend, colour = Source),
                 linewidth = 1.6, lineend = "round") +
    annotate("point", x = 0, y = 0, size = 1.5) +
    geom_text(data = compass, aes(x, y, label = label), size = fig_sizes$annotation) +
    annotate("text", x = 0, y = 1.42, size = fig_sizes$annotation,
             label = sprintf("Mooring: %.0f%% | Model: %.0f%%",
                             pc$Variance_explained_pct_Mooring, pc$Variance_explained_pct_Model)) +
    annotate("text", x = 0, y = -1.42, size = fig_sizes$annotation,
             label = sprintf("Δθ = %.1f°", pc$Axis_difference_deg)) +
    scale_colour_manual(values = source_cols, guide = "none") +
    coord_fixed(xlim = c(-1.3, 1.3), ylim = c(-1.55, 1.55), clip = "off") +
    labs(title = paste0(cy, " ", loc, " ", dep, " m")) +
    theme_void() +
    theme(text = element_text(family = font_family),
          plot.title = element_text(face = "bold", size = fig_sizes$strip, hjust = 0.5),
          plot.tag   = element_text(size = fig_sizes$tag, face = "bold"))
}

s1_panels <- pmap(s1_order, function(Canyon, Location, Depth_m) axis_panel(Canyon, Location, Depth_m))

legend_s1 <- legend_panel(
  ggplot(tibble(Source = factor(names(source_cols), levels = names(source_cols))),
         aes(x = 0, y = 0, colour = Source)) +
    geom_path(alpha = 0, linewidth = 1.6) +
    scale_colour_manual(values = source_cols, name = NULL,
                        guide = guide_legend(override.aes = list(alpha = 1, linewidth = 2.5))),
  scale = 1.4
)

design_s1 <- "
ABCDI
EFGHI
"

figureS2 <- wrap_plots(s1_panels) + legend_s1 +
  plot_layout(design = design_s1, widths = c(1, 1, 1, 1, 0.6)) +
  plot_annotation(tag_levels = list(c(LETTERS[1:8], ""))) &
  tag_inside(0.02, 0.99)

figureS2

save_figure(figureS2, "Figure S2", width = 16, height = 9)
```

# 7. Figure S3 – full-year modelled near-bottom currents

Stick plots of the model currents at all deep stations (daily means):
one row per depth and two columns per canyon (inside / outside). The
stations of each canyon share the y axis. Each
arrow starts at (time, 0) and points to (time + u × `x_scale_factor`, v).
The red line marks the mean current speed (drawn at −mean speed); the
background colours mark the seasons.

```{r figS3, fig.width=26, fig.height=11}
stick_model <- model %>%
  mutate(time = floor_date(time, paste(stick_bin_hours, "hours"))) %>%
  group_by(Canyon, Location, Depth_m, time) %>%
  summarise(u = mean(u, na.rm = TRUE), v = mean(v, na.rm = TRUE), .groups = "drop")

mean_speed_model <- model %>%
  group_by(Canyon, Location, Depth_m) %>%
  summarise(avg_speed = mean(speed, na.rm = TRUE), .groups = "drop")

season_bands <- make_season_bands(model$time)

# y range of each canyon (shared by its In and Out stations and all depths)
ylim_canyon <- stick_model %>%
  left_join(mean_speed_model, by = c("Canyon", "Location", "Depth_m")) %>%
  group_by(Canyon) %>%
  summarise(lo = min(c(v, -avg_speed)), hi = max(v), .groups = "drop")

# Figure S3 is very wide, so its text is enlarged (factor s2_text_scale)
s2_text_scale <- 1.35

theme_stick <- theme_bw() + theme_fig +
  theme(axis.text   = element_text(size = fig_sizes$axis_text * s2_text_scale, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.text  = element_text(size = fig_sizes$strip * 1.15),   # longer station names
        panel.grid.minor = element_blank())

stick_panel_model <- function(cy, loc, dep, show_x, show_y) {
  d  <- filter(stick_model, Canyon == cy, Location == loc, Depth_m == dep) %>%
    mutate(panel = paste0(cy, " ", dep, " m - ", loc))
  ms <- filter(mean_speed_model, Canyon == cy, Location == loc, Depth_m == dep)
  yl <- filter(ylim_canyon, Canyon == cy)

  p <- ggplot(d) +
    geom_rect(data = season_bands, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf,
                                       fill = season),
              inherit.aes = FALSE, alpha = 0.35) +
    geom_hline(yintercept = 0, colour = "grey40", linetype = "dashed", linewidth = 0.7) +
    geom_segment(aes(x = time, y = 0, xend = time + u * x_scale_factor, yend = v),
                 colour = "mediumblue", linewidth = 0.45) +
    geom_hline(yintercept = -ms$avg_speed, colour = "red", linewidth = 0.7, alpha = 0.85) +
    facet_wrap(~ panel) +
    scale_fill_manual(values = season_cols, guide = "none") +
    scale_x_datetime(date_breaks = "2 months", date_labels = "%m/%y",
                     expand = expansion(mult = 0.005)) +
    scale_y_continuous(labels = label_2dec) +
    coord_cartesian(ylim = c(yl$lo, yl$hi)) +
    labs(x = NULL, y = NULL) +
    theme_stick
  # all rows share the same dates: x numbers only under the bottom row
  # (removed, not just hidden, so that the rows sit close together)
  if (!show_x) p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  if (!show_y) p <- p + hide_axis_text("y")
  p
}

# 3 rows (depth) x 6 columns (In / Out of each canyon); y numbers only on the
# "In" column of each canyon, x numbers only under the bottom row
s2_order <- expand_grid(Depth_m = c(100, 200, 300), Canyon = canyon_levels,
                        Location = c("In", "Out"))

s2_panels <- pmap(s2_order, function(Depth_m, Canyon, Location) {
  stick_panel_model(Canyon, Location, Depth_m,
                    show_x = Depth_m == 300,
                    show_y = Location == "In")
})

legend_s2 <- legend_panel(
  ggplot(season_bands, aes(x = 0, y = 0, fill = season)) +
    geom_tile(alpha = 0) +
    scale_fill_manual(values = season_cols, name = "Season:",
                      guide = guide_legend(override.aes = list(alpha = 0.6, colour = "black"))),
  scale = s2_text_scale
)

figureS3_panels <- wrap_plots(s2_panels, ncol = 6) | legend_s2
figureS3_panels <- figureS3_panels + plot_layout(widths = c(1, 0.1)) +
  plot_annotation(tag_levels = list(c(LETTERS[1:18], ""))) &
  tag_inside(0.01, 0.99)

figureS3 <- add_figure_titles(figureS3_panels, y = "North–South velocity (cm s⁻¹)",
                              size = fig_sizes$axis_title * s2_text_scale)

figureS3

save_figure(figureS3, "Figure S3", width = 26, height = 11)
```

# Save the workspace and the R session information

```{r save-session}
save.image(file = file.path(out_dir, "02_Model_Mooring_Comparison.RData"))
sessionInfo()
```
