---
title: "Script 3 – Fish community and environment (Figures 6, 7, 8, S4, S5)"
author: "Almog Haim – MEDsLab, Bar-Ilan University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

# Overview

This script analyses the BRUVS fish data together with the environmental data
of each deployment, starting from the two raw tables, and reproduces:

| Figure | Content |
|---|---|
| 6  | Sample-based species rarefaction curves, inside vs. outside the canyons |
| 7  | GAMs – species richness and total MaxN by location (In / Out) |
| 8  | GAMs – species richness and total MaxN by canyon |
| S4 | Substrate type by depth range in each canyon |
| S5 | Redundancy analysis (RDA) of the fish community |

The script also runs the model diagnostics reported in the article
(`gam.check()` and DHARMa residual tests, Section 4.3).

Code consolidated from the original BRUVS analysis script.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)   # dplyr, tidyr, ggplot2, readr, stringr, purrr
library(vegan)       # specaccum, decostand, rda
library(mgcv)        # GAMs
library(DHARMa)      # simulation-based residual diagnostics of the GAMs
library(ggrepel)     # non-overlapping labels in the RDA biplot
library(patchwork)   # multi-panel figure assembly
```

# 1. Input files and settings

All paths are relative to this script (the `scripts` folder); the input
files are in `data/bruvs` and all outputs are written to `outputs`.

```{r paths}
data_dir    <- file.path("..", "data", "bruvs")
fish_file   <- file.path(data_dir, "bruvs_maxn.csv")          # one row per taxon per deployment
deploy_file <- file.path(data_dir, "bruvs_deployments.csv")   # one row per deployment

# Output folder for figures, tables and the saved workspace
out_dir <- file.path("..", "outputs")
fig_dir <- out_dir
dir.create(fig_dir, showWarnings = FALSE)

missing <- c(fish_file, deploy_file)[!file.exists(c(fish_file, deploy_file))]
if (length(missing) > 0) stop("Missing input file(s):\n", paste(missing, collapse = "\n"))
```

```{r settings}
canyon_levels   <- c("Eilat", "Shlomo", "Princess")
location_levels <- c("In", "Out")

canyon_cols   <- c("Eilat" = "steelblue", "Shlomo" = "darkorange", "Princess" = "darkgreen")
location_cols <- c("In" = "#d95f02", "Out" = "#7570b3")

substrate_bin  <- 20    # depth bins of Figure S4 (m)
rda_arrow_scale <- 1.4  # stretch of the arrows in the RDA biplot

# ---------------------------------------------------------------------------
# Figure style – used by every figure so that all figures look the same
# (identical to Scripts 1 and 2)
# ---------------------------------------------------------------------------
# Font: Times New Roman, requested through R's generic "serif" family.
# "serif" is understood by every R graphics device (screen, png, pdf,
# cairo_pdf, ragg) and is drawn as Times New Roman on Windows (Times on
# macOS). Asking for "Times New Roman" by name makes the standard pdf() and
# png() devices fail silently, which produces empty files.
font_family <- "serif"

# Text sizes (pt) shared by all figures
fig_sizes <- list(
  axis_title   = 18,   # axis titles
  axis_text    = 14,   # axis numbers
  strip        = 16,   # panel title strips
  legend_title = 16,
  legend_text  = 14,
  legend_key   = 1.3,  # size of the legend colour boxes (lines)
  tag          = 18,   # panel letters
  annotation   = 4.5   # text inside panels (mm, ggplot units)
)

theme_fig <- theme(
  text         = element_text(family = font_family),
  axis.title   = element_text(size = fig_sizes$axis_title),
  axis.text    = element_text(size = fig_sizes$axis_text, colour = "black"),
  strip.text   = element_text(size = fig_sizes$strip),
  legend.title = element_text(size = fig_sizes$legend_title, face = "bold"),
  legend.text  = element_text(size = fig_sizes$legend_text),
  plot.tag     = element_text(size = fig_sizes$tag, face = "bold")
)

# Text drawn inside the panels (geom_text / geom_label) uses the same font
update_geom_defaults("text",  list(family = font_family))
update_geom_defaults("label", list(family = font_family))

# Single y-axis title (left) and x-axis title (bottom) for a whole figure
add_figure_titles <- function(panels, y = NULL, x = NULL, x_hjust = 0.5,
                              size = fig_sizes$axis_title) {
  wrap_elements(panel = panels) +
    labs(tag = y, caption = x) +
    theme(text              = element_text(family = font_family),
          plot.tag          = element_text(angle = 90, size = size),
          plot.tag.position = "left",
          plot.caption      = element_text(size = size, hjust = x_hjust))
}

# Hide axis numbers but keep their space, so that panels stay evenly spaced
hide_axis_text <- function(axis = c("x", "y")) {
  axis <- match.arg(axis)
  if (axis == "x") {
    theme(axis.text.x = element_text(colour = "transparent"),
          axis.ticks.x = element_line(colour = "transparent"))
  } else {
    theme(axis.text.y = element_text(colour = "transparent"),
          axis.ticks.y = element_line(colour = "transparent"))
  }
}

# Save a figure as PDF (vector, embedded font), 300-dpi PNG and 300-dpi TIFF.
# Always pass the figure object explicitly: ggsave() without `plot =` saves
# ggplot2's "last plot", which for a patchwork figure can be one of its
# internal pieces (e.g. the empty legend panel).
save_figure <- function(plot, name, width, height, dpi = 300) {
  has_ragg   <- requireNamespace("ragg", quietly = TRUE)
  pdf_device <- if (capabilities("cairo")) cairo_pdf else "pdf"
  files <- file.path(fig_dir, paste0(name, c(".pdf", ".png", ".tiff")))
  start <- Sys.time() - 1

  ggsave(files[1], plot, width = width, height = height, device = pdf_device)
  ggsave(files[2], plot, width = width, height = height, dpi = dpi, bg = "white",
         device = if (has_ragg) ragg::agg_png else "png")
  ggsave(files[3], plot, width = width, height = height, dpi = dpi, bg = "white",
         device = if (has_ragg) ragg::agg_tiff else "tiff", compression = "lzw")

  # Windows cannot overwrite a file that is open in another program (image
  # viewer, PDF reader, Explorer preview pane). Check that every file was
  # really rewritten and say so if not.
  not_updated <- files[is.na(file.mtime(files)) | file.mtime(files) < start]
  if (length(not_updated) > 0) {
    warning("Not updated (probably open in another program - close it and run again):\n  ",
            paste(not_updated, collapse = "\n  "), call. = FALSE)
  } else {
    message("Saved: ", paste(basename(files), collapse = ", "))
  }

  set_last_plot(plot)   # so that a later ggsave() without `plot =` also saves this figure
  invisible(plot)
}

# Panel letters in the top-left corner of each panel
tag_inside <- function(x = 0.01, y = 0.99, size = fig_sizes$tag) {
  theme(plot.tag = element_text(size = size, face = "bold"), plot.tag.position = c(x, y))
}

# Axis labels with at most two decimals, trailing zeros dropped
label_2dec  <- function(x) sub("\\.?0+$", "", sprintf("%.2f", x))

# Legend drawn from a separate, otherwise empty plot, so that it appears once
# and looks the same in every ggplot2 version
# `scale` enlarges text and keys (e.g. 1.25 = 25 % larger than the default)
legend_panel <- function(p, scale = 1) {
  wrap_elements(full = p + theme_void() +
    theme(legend.position = "left",
          legend.key.size = unit(fig_sizes$legend_key * scale, "lines"),
          text         = element_text(family = font_family),
          legend.title = element_text(face = "bold", size = fig_sizes$legend_title * scale),
          legend.text  = element_text(size = fig_sizes$legend_text * scale)))
}

theme_panel <- theme_classic() + theme_fig
```

# 2. Load and clean the data

## 2.1 Fish observations

```{r load-fish}
fish_raw   <- read.csv(fish_file, check.names = FALSE)
deploy_raw <- read.csv(deploy_file, check.names = FALSE)

# Text clean-up: some species names contain non-breaking spaces
# (e.g. "Choerodon robustus" typed twice, once with each kind of space),
# which would turn one species into two different columns.
clean_text <- function(x) na_if(str_squish(gsub(" ", " ", x)), "")

fish <- fish_raw %>%
  mutate(across(where(is.character), ~ iconv(.x, from = "", to = "UTF-8", sub = "")),
         across(c(OpCode, Canyon, Location, Family, Species), clean_text)) %>%
  filter(!is.na(OpCode), !is.na(MaxN), MaxN >= 0)   # empty deployments have no fish rows
```

## 2.2 Inside / outside the canyon (In / Out) of every deployment

`Deploy_summary.csv` (column `Canyon_Location`) and `red_MaxN_data.csv`
(column `Location`) both give the position of the deployment. When they
disagree, the value in `red_MaxN_data.csv` is used. Positions can also be
set by hand in `location_manual` (these override both files); it is empty,
because the two files were checked and corrected and now agree. The script stops if a deployment is missing from one of the files
or has more than one position in `red_MaxN_data.csv`.

```{r location}
# Manual corrections (none: both files were corrected and agree)
location_manual <- tibble(OpCode = character(), Location_manual = character())

location_check <- full_join(
  deploy_raw %>% transmute(OpCode = clean_text(OpCode), Location_deploy = clean_text(Canyon_Location)),
  fish_raw   %>% transmute(OpCode = clean_text(OpCode), Location_fish   = clean_text(Location)) %>%
    filter(!is.na(OpCode)) %>% distinct(),
  by = "OpCode")

unclear <- location_check %>%
  filter(is.na(Location_deploy) | is.na(Location_fish) |
         OpCode %in% OpCode[duplicated(OpCode)])
if (nrow(unclear) > 0) {
  print(unclear)
  stop("Unclear In/Out position for the deployments above - check the two files")
}
stopifnot(all(location_manual$OpCode %in% location_check$OpCode))

location_table <- location_check %>%
  left_join(location_manual, by = "OpCode") %>%
  mutate(Location = coalesce(Location_manual, Location_fish),
         Source   = case_when(!is.na(Location_manual)        ~ "manual correction",
                              Location_deploy != Location_fish ~ "red_MaxN_data.csv (files disagree)",
                              TRUE                             ~ "both files agree"))

# Deployments where the two files disagree or a manual correction was used
location_table %>% filter(Source != "both files agree") %>% knitr::kable()

fish <- fish %>% select(-Location) %>% left_join(select(location_table, OpCode, Location), by = "OpCode")
```

## 2.3 One row per deployment

Species richness = number of different taxa recorded in the deployment;
total MaxN = sum of the MaxN of all taxa. Deployments without fish get 0.

```{r deployments}
fish_summary <- fish %>%
  group_by(OpCode) %>%
  summarise(Richness   = n_distinct(Species[MaxN > 0]),
            Total_MaxN = sum(MaxN),
            .groups = "drop")

deployments <- deploy_raw %>%
  transmute(OpCode = clean_text(OpCode),
            Canyon = clean_text(Canyon),
            depth       = suppressWarnings(as.numeric(depth)),
            temperature = suppressWarnings(as.numeric(temperature)),
            Substrate   = na_if(clean_text(Substrate), "na")) %>%
  left_join(select(location_table, OpCode, Location), by = "OpCode") %>%
  left_join(fish_summary, by = "OpCode") %>%
  mutate(Richness   = replace_na(Richness, 0L),
         Total_MaxN = replace_na(Total_MaxN, 0L),
         Canyon     = factor(Canyon, levels = canyon_levels),
         Location   = factor(Location, levels = location_levels),
         Substrate  = factor(Substrate, levels = c("Hard", "Mixed", "Soft")))

stopifnot(!anyNA(deployments$Canyon), !anyNA(deployments$Location))

# Number of deployments
deployments %>% count(Canyon, Location) %>%
  pivot_wider(names_from = Location, values_from = n) %>% knitr::kable()
```

## 2.4 Community matrix

Presence (1) / absence (0) of every species (records identified to species
level) in every deployment, including the deployments without fish.

```{r community}
community <- fish %>%
  filter(!is.na(Species), MaxN > 0) %>%
  distinct(OpCode, Species) %>%
  mutate(present = 1) %>%
  pivot_wider(names_from = Species, values_from = present, values_fill = 0)

community <- deployments %>% select(OpCode) %>%
  left_join(community, by = "OpCode") %>%
  mutate(across(-OpCode, ~ replace_na(.x, 0))) %>%
  column_to_rownames("OpCode") %>%
  as.matrix()

stopifnot(identical(rownames(community), deployments$OpCode))
cat(ncol(community), "species in", nrow(community), "deployments\n")
```

# 3. Figure 6 – sample-based species rarefaction curves

Sample-based rarefaction curves (`vegan::specaccum(method = "exact")`): the
expected number of species in a random subset of *m* deployments, computed
analytically from the species incidences (presence/absence in each deployment)
instead of averaging random orderings of the deployments (Colwell et al. 2004,
Ecology 85: 2717–2727). Shading: ± 1.96 SD of the expected richness.
The SD is unconditional (`conditioned = FALSE`): the observed deployments are
treated as a sample from a larger assemblage, whose total richness is estimated
with the Chao2 estimator (`gamma = "chao"`); therefore the SD does not shrink to
zero at the full number of deployments.

```{r fig6, fig.width=15, fig.height=6.5}
rarefaction <- function(rows, ...) {
  rc <- specaccum(community[rows, , drop = FALSE], method = "exact",
                conditioned = FALSE, gamma = "chao")
  tibble(Samples = rc$sites, Richness = rc$richness, SD = rc$sd, ...)
}

rare_all <- map_dfr(location_levels, function(l)
  rarefaction(deployments$Location == l, Location = l))

rare_canyon <- expand_grid(Canyon = canyon_levels, Location = location_levels) %>%
  pmap_dfr(function(Canyon, Location)
    rarefaction(deployments$Canyon == Canyon & deployments$Location == Location,
                 Canyon = Canyon, Location = Location))

# Number of deployments in each curve, written in the panel titles
n_label <- function(d) paste0("In = ", sum(d$Location == "In"), ", Out = ", sum(d$Location == "Out"))
strip_all    <- paste0("All canyons\n", n_label(deployments))
strip_canyon <- deployments %>% group_by(Canyon) %>%
  summarise(label = paste0(first(Canyon), "\nIn = ", sum(Location == "In"),
                           ", Out = ", sum(Location == "Out")), .groups = "drop")

rare_all    <- rare_all %>% mutate(Panel = strip_all)
rare_canyon <- rare_canyon %>%
  mutate(Canyon = factor(Canyon, levels = canyon_levels)) %>%
  left_join(strip_canyon, by = "Canyon") %>%
  mutate(Panel = factor(label, levels = strip_canyon$label))

rarefaction_panel <- function(d) {
  ggplot(d, aes(Samples, Richness, colour = Location, fill = Location)) +
    geom_ribbon(aes(ymin = pmax(0, Richness - 1.96 * SD), ymax = Richness + 1.96 * SD),
                alpha = 0.18, colour = NA) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~ Panel, nrow = 1) +
    scale_colour_manual(values = location_cols, guide = "none") +
    scale_fill_manual(values = location_cols, guide = "none") +
    scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
    labs(x = NULL, y = NULL) +
    theme_panel +
    theme(strip.background = element_rect(colour = "black", linewidth = 0.6))
}

p6A <- rarefaction_panel(rare_all)
p6B <- rarefaction_panel(rare_canyon) +
  scale_x_continuous(breaks = seq(0, 20, by = 4))

legend_location <- legend_panel(
  ggplot(tibble(Location = factor(location_levels, levels = location_levels), x = 0, y = 0),
         aes(x, y, colour = Location, fill = Location)) +
    geom_ribbon(aes(ymin = y, ymax = y), alpha = 0.18, colour = NA) +
    geom_line(linewidth = 1.2, alpha = 0) +
    scale_colour_manual(values = location_cols, name = "Location:",
                        guide = guide_legend(override.aes = list(alpha = 1))) +
    scale_fill_manual(values = location_cols, name = "Location:"),
  scale = 1.2
)

figure6_panels <- p6A + p6B + legend_location +
  plot_layout(widths = c(1, 2.1, 0.32)) +
  plot_annotation(tag_levels = list(c("A", "B", ""))) &
  tag_inside(0.005, 0.995)

figure6 <- add_figure_titles(figure6_panels, y = "Expected species richness",
                             x = "Number of deployments", x_hjust = 0.42)
figure6

save_figure(figure6, "Figure 6", width = 15, height = 6.5)
```

# 4. GAMs – species richness and total MaxN

Species richness: Poisson GAM. Total MaxN: negative-binomial GAM (over-dispersed
counts). All candidate models include canyon and location as fixed effects and
smooth effects of depth and near-bottom temperature; they differ in whether
the smooths are shared by all deployments or fitted separately for each
location (`by = Location`) or each canyon (`by = Canyon`).

## 4.1 Model selection (AIC)

Every candidate model is fitted to the same data and compared with `AIC()`,
together with a null model (intercept only, `model_*_0`) that shows whether
the explanatory variables explain the response at all. The models used in
the article are set in `use_richness_model` / `use_maxn_model` (the
lowest-AIC candidate models); the script warns if another candidate model
has a lower AIC.

Note on `R-sq.(adj)` in `summary()`: for count models mgcv computes it from
the raw residuals (observed − fitted), so a few very large counts can make it
negative. For Poisson / negative-binomial GAMs the deviance explained is the
usual measure; both are given in the AIC table.

```{r model-selection-richness}
# --- Species richness (Poisson) ---------------------------------------------
model_richness_1 <- gam(
  Richness ~ Canyon + Location +
    s(depth, k = 5) +
    s(temperature, k = 5),
  family = poisson(), data = deployments, method = "REML")

model_richness_2 <- gam(
  Richness ~ Canyon + Location +
    s(depth, by = Canyon, k = 5) +
    s(temperature, by = Canyon, k = 5),
  family = poisson(), data = deployments, method = "REML")

model_richness_3 <- gam(
  Richness ~ Canyon + Location +
    s(depth, by = Location, k = 5) +
    s(temperature, by = Location, k = 5),
  family = poisson(), data = deployments, method = "REML")

model_richness_0 <- gam(Richness ~ 1, family = poisson(), data = deployments, method = "REML")

# AIC of the species-richness models
aic_richness <- AIC(model_richness_0, model_richness_1, model_richness_2, model_richness_3)
aic_richness
```

```{r model-selection-maxn}
# --- Total MaxN (negative binomial) -----------------------------------------
model_maxn_1 <- gam(
  Total_MaxN ~ Canyon + Location +
    s(depth, by = Canyon, k = 5) +
    s(temperature, by = Canyon, k = 5),
  family = nb(), data = deployments, method = "REML")

model_maxn_2 <- gam(
  Total_MaxN ~ Canyon + Location +
    s(depth, by = Location, k = 5) +
    s(temperature, k = 5),
  family = nb(), data = deployments, method = "REML")

model_maxn_3 <- gam(
  Total_MaxN ~ Canyon + Location +
    s(depth, by = Location, k = 5) +
    s(temperature, by = Location, k = 5),
  family = nb(), data = deployments, method = "REML")

model_maxn_4 <- gam(
  Total_MaxN ~ Canyon + Location +
    s(depth, k = 5) +
    s(temperature, k = 5),
  family = nb(), data = deployments, method = "REML")

model_maxn_0 <- gam(Total_MaxN ~ 1, family = nb(), data = deployments, method = "REML")
# AIC of the total-MaxN models
aic_maxn <- AIC(model_maxn_0, model_maxn_1, model_maxn_2, model_maxn_3, model_maxn_4)
aic_maxn
```

```{r model-selection-table}
# Models chosen for the article (the lowest-AIC candidate models). Set to NULL
# to use the lowest-AIC candidate model automatically.
use_richness_model <- "model_richness_3"
use_maxn_model     <- "model_maxn_4"

# AIC table with delta AIC, Akaike weight, deviance explained and R-sq.(adj)
aic_summary <- function(aic, response, env = parent.frame()) {
  as_tibble(aic, rownames = "Model") %>%
    mutate(Response = response,
           Formula  = map_chr(Model, ~ paste(deparse(formula(get(.x, envir = env)), width.cutoff = 500), collapse = "")),
           Deviance_explained = map_dbl(Model, ~ summary(get(.x, envir = env))$dev.expl),
           R_sq_adj = map_dbl(Model, ~ summary(get(.x, envir = env))$r.sq),
           Delta_AIC  = AIC - min(AIC),
           AIC_weight = exp(-Delta_AIC / 2) / sum(exp(-Delta_AIC / 2))) %>%
    relocate(Response, Model) %>%
    arrange(AIC)
}

aic_table <- bind_rows(aic_summary(aic_richness, "Species richness"),
                       aic_summary(aic_maxn,     "Total MaxN"))
aic_table %>% select(-Formula) %>% knitr::kable(digits = 3)
write_csv(aic_table, file.path(fig_dir, "Table_GAM_AIC.csv"))

# Lowest-AIC candidate model (the null models are not candidates)
best_candidate <- function(aic) {
  aic <- aic[!grepl("_0$", rownames(aic)), ]
  rownames(aic)[which.min(aic$AIC)]
}
selected_richness <- use_richness_model %||% best_candidate(aic_richness)
selected_maxn     <- use_maxn_model     %||% best_candidate(aic_maxn)

for (sel in list(c(selected_richness, best_candidate(aic_richness)),
                 c(selected_maxn,     best_candidate(aic_maxn))))
  if (sel[1] != sel[2]) warning(sel[1], " is used, but ", sel[2], " has a lower AIC", call. = FALSE)

# Is the selected model better than the null model?
for (x in list(list(aic_richness, selected_richness, "model_richness_0"),
               list(aic_maxn,     selected_maxn,     "model_maxn_0"))) {
  d_aic <- x[[1]][x[[2]], "AIC"] - x[[1]][x[[3]], "AIC"]
  cat(sprintf("%s: AIC %.1f %s the null model\n", x[[2]], abs(d_aic),
              if (d_aic < 0) "lower (better) than" else "HIGHER (worse) than"))
}

model_richness <- get(selected_richness)
model_maxn     <- get(selected_maxn)

cat("Species richness model used:", selected_richness, "\n ", deparse(formula(model_richness), 500), "\n")
cat("Total MaxN model used:      ", selected_maxn,     "\n ", deparse(formula(model_maxn), 500), "\n")
```

## 4.2 Selected models

```{r gams}
summary(model_richness)
concurvity(model_richness)
summary(model_maxn)
concurvity(model_maxn)
```

## 4.3 Model diagnostics

Diagnostics of the two models used in the article (Methods: "Model adequacy
was evaluated through inspection of convergence, residual patterns, effective
degrees of freedom, and the k-index using gam.check() … Simulation-based
residual diagnostics were additionally performed using the DHARMa package …
including tests for dispersion, residual uniformity, zero inflation, and
outliers").

**mgcv checks.** `gam.check()` prints the convergence of the REML
optimisation and, for every smooth, the basis dimension (k'), the effective
degrees of freedom (edf) and the k-index with its p-value (a low k-index with
a small p-value suggests that k is too small), and draws the standard residual
plots. The same numbers are collected in a table with `k.check()`.

```{r diagnostics-mgcv, fig.width=9, fig.height=8}
set.seed(5)   # the k-index p-values are based on random permutations

cat("--- Species richness model:", selected_richness, "---\n")
par(mfrow = c(2, 2))
gam.check(model_richness, rep = 500)

cat("\n--- Total MaxN model:", selected_maxn, "---\n")
par(mfrow = c(2, 2))
gam.check(model_maxn, rep = 500)
par(mfrow = c(1, 1))

# Convergence and basis-dimension checks in one table
k_table <- function(model, response) {
  as_tibble(k.check(model), rownames = "Term") %>%
    transmute(Response = response, Term,
              k_prime = `k'`, edf, k_index = `k-index`, p_value = `p-value`,
              Converged = model$converged,
              REML_convergence = model$outer.info$conv)
}
diagnostics_k <- bind_rows(k_table(model_richness, "Species richness"),
                           k_table(model_maxn,     "Total MaxN"))
diagnostics_k %>% knitr::kable(digits = 3)
```

**DHARMa checks.** Scaled (quantile) residuals are obtained by simulating
new response values from each fitted model (1000 simulations): Poisson for
species richness and negative binomial, with the estimated dispersion
parameter (theta), for total MaxN. The simulations are drawn directly from the
fitted means, so that the same procedure is used for both families, and are
passed to DHARMa with `createDHARMa()`. If a model fits well, the scaled
residuals are uniformly distributed between 0 and 1. The tests are:
dispersion (`testDispersion`), uniformity of the residuals
(`testUniformity`, Kolmogorov–Smirnov), excess zeros (`testZeroInflation`)
and outliers (`testOutliers`, bootstrap). The residuals are also plotted
against the fitted values and against every predictor.

```{r diagnostics-dharma, fig.width=10, fig.height=5}
n_sim <- 1000

# Simulated responses (one column per simulation) from a Poisson or
# negative-binomial GAM
simulate_gam <- function(model, n_sim) {
  mu  <- fitted(model)
  fam <- model$family$family
  if (fam == "poisson") {
    replicate(n_sim, rpois(length(mu), mu))
  } else if (startsWith(fam, "Negative Binomial")) {
    theta <- model$family$getTheta(TRUE)
    replicate(n_sim, rnbinom(length(mu), mu = mu, size = theta))
  } else {
    stop("simulate_gam(): family not supported: ", fam)
  }
}

dharma_residuals <- function(model, seed) {
  set.seed(seed)
  createDHARMa(simulatedResponse      = simulate_gam(model, n_sim),
               observedResponse       = model$y,
               fittedPredictedResponse = fitted(model),
               integerResponse        = TRUE)
}

res_richness <- dharma_residuals(model_richness, seed = 6)
res_maxn     <- dharma_residuals(model_maxn,     seed = 7)

# Residual plots: QQ plot with tests (left) and residuals vs. fitted values (right),
# then residuals against each predictor (data of the model, same row order)
residual_plots <- function(res, model, title) {
  plot(res, title = title)
  mf <- model$model
  par(mfrow = c(2, 2))
  plotResiduals(res, form = mf$depth,       xlab = "Depth (m)",       main = "Depth")
  plotResiduals(res, form = mf$temperature, xlab = "Temperature (°C)", main = "Temperature")
  plotResiduals(res, form = mf$Location,    xlab = "Location",        main = "Location")
  plotResiduals(res, form = mf$Canyon,      xlab = "Canyon",          main = "Canyon")
  par(mfrow = c(1, 1))
}
residual_plots(res_richness, model_richness, "Species richness")
residual_plots(res_maxn,     model_maxn,     "Total MaxN")

# Test results in one table
dharma_tests <- function(res, response) {
  tests <- list(
    "Dispersion"           = testDispersion(res, plot = FALSE),
    "Uniformity (KS)"      = testUniformity(res, plot = FALSE),
    "Zero inflation"       = testZeroInflation(res, plot = FALSE),
    "Outliers (bootstrap)" = testOutliers(res, type = "bootstrap", plot = FALSE)
  )
  tibble(Response  = response,
         Test      = names(tests),
         Statistic = map_dbl(tests, ~ unname(.x$statistic[1])),
         p_value   = map_dbl(tests, "p.value"))
}
diagnostics_dharma <- bind_rows(dharma_tests(res_richness, "Species richness"),
                                dharma_tests(res_maxn,     "Total MaxN"))
diagnostics_dharma %>% knitr::kable(digits = 3)

write_csv(diagnostics_k,      file.path(fig_dir, "Table_GAM_diagnostics_k.csv"))
write_csv(diagnostics_dharma, file.path(fig_dir, "Table_GAM_diagnostics_DHARMa.csv"))
```

```

## 4.4 Model predictions

```{r prediction-functions}
# Model predictions on the response scale, with 95 % confidence intervals.
# type = "average":  prediction averaged over all deployments, with the focal
#                    variables set to the given values (= ggeffects
#                    predict_response(margin = "empirical")); delta-method CI.
# type = "marginal": prediction at the mean of the other numeric variable,
#                    averaged with equal weights over the levels of the other
#                    factor(s) on the link scale (= ggeffects
#                    predict_response(margin = "marginalmeans", weights =
#                    "equal")); CI from the t distribution with the residual
#                    degrees of freedom, back-transformed.
# Both reproduce the values of the original script (ggeffects) exactly.
model_predictions <- function(model, data, x_var, x_values, by_var,
                              by_levels = levels(data[[by_var]]),
                              type = c("average", "marginal")) {
  type <- match.arg(type)
  b <- coef(model); V <- vcov(model)
  inv <- model$family$linkinv; mu_eta <- model$family$mu.eta
  vars    <- all.vars(formula(model))[-1]
  factors <- vars[sapply(data[vars], is.factor)]
  numeric <- setdiff(vars, factors)

  if (type == "marginal") {
    other_f <- setdiff(factors, by_var)
    base <- expand_grid(!!!lapply(set_names(other_f), function(v) levels(data[[v]])))
    for (v in setdiff(numeric, x_var)) base[[v]] <- mean(data[[v]], na.rm = TRUE)
    for (v in other_f) base[[v]] <- factor(base[[v]], levels = levels(data[[v]]))
  }

  map_dfr(by_levels, function(g) map_dfr(x_values, function(x) {
    nd <- if (type == "average") data else base
    nd[[x_var]]  <- x
    nd[[by_var]] <- factor(g, levels = levels(data[[by_var]]))
    X <- predict(model, newdata = nd, type = "lpmatrix")
    if (type == "average") {
      eta  <- drop(X %*% b)
      est  <- mean(inv(eta))
      grad <- colMeans(X * mu_eta(eta))
      se   <- sqrt(drop(grad %*% V %*% grad))
      tibble(x = x, group = g, predicted = est,
             conf.low = est - 1.96 * se, conf.high = est + 1.96 * se)
    } else {
      xbar <- colMeans(X)
      eta  <- sum(xbar * b)
      se   <- sqrt(drop(xbar %*% V %*% xbar))
      q    <- qt(0.975, df.residual(model))
      tibble(x = x, group = g, predicted = inv(eta),
             conf.low = inv(eta - q * se), conf.high = inv(eta + q * se))
    }
  })) %>% mutate(group = factor(group, levels = by_levels))
}

# Difference in the average prediction between each level of `variable` and
# the reference level, with 95 % CI (= marginaleffects::avg_comparisons(type =
# "response"))
level_differences <- function(model, data, variable,
                              reference = levels(data[[variable]])[1]) {
  b <- coef(model); V <- vcov(model)
  inv <- model$family$linkinv; mu_eta <- model$family$mu.eta
  lv <- levels(data[[variable]])
  average_at <- function(level) {
    nd <- data; nd[[variable]] <- factor(level, levels = lv)
    X <- predict(model, newdata = nd, type = "lpmatrix"); eta <- drop(X %*% b)
    list(est = mean(inv(eta)), grad = colMeans(X * mu_eta(eta)))
  }
  ref <- average_at(reference)
  map_dfr(lv, function(l) {
    a <- average_at(l); d <- a$est - ref$est; g <- a$grad - ref$grad
    se <- sqrt(drop(g %*% V %*% g))
    tibble(level = l, estimate = d, conf.low = d - 1.96 * se, conf.high = d + 1.96 * se)
  }) %>%
    mutate(across(c(estimate, conf.low, conf.high), ~ if_else(level == reference, 0, .x)),
           level = factor(level, levels = lv))
}
```

```{r predictions}
d <- deployments
grid200 <- function(x) seq(min(x, na.rm = TRUE), max(x, na.rm = TRUE), length.out = 200)

# --- Species richness: all observed depths / temperatures (as in the original)
rich_depth_location <- model_predictions(model_richness, d, "depth", sort(unique(d$depth)), "Location")
rich_temp_location  <- model_predictions(model_richness, d, "temperature", sort(unique(d$temperature)), "Location")
rich_depth_canyon   <- model_predictions(model_richness, d, "depth", sort(unique(d$depth)), "Canyon")
rich_temp_canyon    <- model_predictions(model_richness, d, "temperature", sort(unique(d$temperature)), "Canyon")

# --- Total MaxN: depth limited to the range sampled by each group
#     by location: the depth range of the In and of the Out deployments
maxn_depth_location <- map_dfr(location_levels, function(l)
  model_predictions(model_maxn, d, "depth", grid200(d$depth[d$Location == l]),
                    "Location", by_levels = l, type = "marginal")) %>%
  mutate(group = factor(group, levels = location_levels))

#     by canyon: the depth range shared by the In and Out deployments of all canyons
depth_overlap <- d %>%
  group_by(Canyon, Location) %>%
  summarise(min_depth = min(depth), max_depth = max(depth), .groups = "drop") %>%
  group_by(Canyon) %>%
  summarise(overlap_min = max(min_depth), overlap_max = min(max_depth), .groups = "drop")
depth_common <- seq(max(depth_overlap$overlap_min), min(depth_overlap$overlap_max), length.out = 200)

maxn_depth_canyon  <- model_predictions(model_maxn, d, "depth", depth_common, "Canyon", type = "marginal")
maxn_temp_location <- model_predictions(model_maxn, d, "temperature", grid200(d$temperature), "Location", type = "marginal")
maxn_temp_canyon   <- model_predictions(model_maxn, d, "temperature", grid200(d$temperature), "Canyon", type = "marginal")

# --- Differences: Out vs. In, and Shlomo / Princess vs. Eilat
diff_rich_location <- level_differences(model_richness, d, "Location", "In")
diff_maxn_location <- level_differences(model_maxn,     d, "Location", "In")
diff_rich_canyon   <- level_differences(model_richness, d, "Canyon",   "Eilat")
diff_maxn_canyon   <- level_differences(model_maxn,     d, "Canyon",   "Eilat")

bind_rows(
  diff_rich_location %>% mutate(Response = "Species richness"),
  diff_maxn_location %>% mutate(Response = "Total MaxN"),
  diff_rich_canyon   %>% mutate(Response = "Species richness"),
  diff_maxn_canyon   %>% mutate(Response = "Total MaxN")
) %>% relocate(Response) %>% knitr::kable(digits = 3)
```

```{r export-tables}
gam_table <- function(model, response) {
  s <- summary(model)
  bind_rows(
    as_tibble(s$p.table, rownames = "Term") %>%
      transmute(Response = response, Term, Type = "parametric",
                Estimate, SE = `Std. Error`, Statistic = `z value`, p = `Pr(>|z|)`),
    as_tibble(s$s.table, rownames = "Term") %>%
      transmute(Response = response, Term, Type = "smooth",
                edf, Ref.df, Statistic = Chi.sq, p = `p-value`)
  ) %>% mutate(Deviance_explained = s$dev.expl, n = s$n)
}

write_csv(bind_rows(gam_table(model_richness, "Species richness"),
                    gam_table(model_maxn, "Total MaxN"),
                    gam_table(maxn_models_sens[[selected_maxn]],
                              "Total MaxN (without extreme counts)")),
          file.path(fig_dir, "Table_GAM_summaries.csv"))

write_csv(bind_rows(
  diff_rich_location %>% mutate(Response = "Species richness", Comparison = "vs In"),
  diff_maxn_location %>% mutate(Response = "Total MaxN",       Comparison = "vs In"),
  diff_rich_canyon   %>% mutate(Response = "Species richness", Comparison = "vs Eilat"),
  diff_maxn_canyon   %>% mutate(Response = "Total MaxN",       Comparison = "vs Eilat")),
  file.path(fig_dir, "Table_GAM_differences.csv"))

write_csv(location_table, file.path(fig_dir, "Table_deployment_location.csv"))
```

## 4.5 Figure panels

```{r gam-panels}
# Predicted response (line + 95 % CI) with the observed deployments (points)
prediction_panel <- function(pred, x_var, y_var, by_var, cols, jitter = FALSE) {
  set.seed(2)   # reproducible jitter of the points
  ggplot() +
    geom_ribbon(data = pred, aes(x = x, ymin = conf.low, ymax = conf.high, fill = group),
                alpha = 0.18, colour = NA) +
    geom_line(data = pred, aes(x = x, y = predicted, colour = group), linewidth = 1.2) +
    geom_point(data = deployments,
               aes(x = .data[[x_var]], y = .data[[y_var]], colour = .data[[by_var]]),
               position = if (jitter) position_jitter(width = 0.03, height = 0) else "identity",
               size = 2.3, alpha = 0.55) +
    scale_colour_manual(values = cols, guide = "none") +
    scale_fill_manual(values = cols, guide = "none") +
    scale_y_continuous(labels = label_2dec) +
    theme_panel
}

# Difference from the reference level (thick bar) with 95 % CI (error bar)
difference_panel <- function(diffs, cols) {
  ggplot(diffs, aes(x = level, colour = level)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.7) +
    geom_errorbar(data = filter(diffs, conf.low != conf.high),
                  aes(ymin = conf.low, ymax = conf.high), width = 0.5, linewidth = 1.1, alpha = 0.6) +
    geom_errorbar(aes(ymin = estimate, ymax = estimate), width = 0.8, linewidth = 2.6) +
    scale_colour_manual(values = cols, guide = "none") +
    scale_x_discrete(limits = levels(diffs$level)) +   # reference level first
    scale_y_continuous(labels = label_2dec) +
    theme_panel
}

# Legend: line, CI band and point, as in the panels
prediction_legend <- function(levels, cols, title) {
  legend_panel(
    ggplot(tibble(g = factor(levels, levels = levels), x = 0, y = 0),
           aes(x, y, colour = g, fill = g)) +
      geom_ribbon(aes(ymin = y, ymax = y), alpha = 0.18, colour = NA) +
      geom_line(linewidth = 1.2, alpha = 0) +
      geom_point(size = 2.3, alpha = 0) +
      scale_colour_manual(values = cols, name = title,
                          guide = guide_legend(override.aes = list(alpha = 1))) +
      scale_fill_manual(values = cols, name = title),
    scale = 1.2)
}

# Six panels (2 rows x 3 columns) + legend; the two prediction panels of a row
# share the y axis, the panels of a column share the x axis
gam_figure <- function(rich_depth, rich_temp, rich_diff, maxn_depth, maxn_temp, maxn_diff,
                       by_var, cols, legend_title, diff_xlab) {
  row_limits <- function(preds, y_var)
    range(0, deployments[[y_var]], unlist(map(preds, ~ c(.x$conf.low, .x$conf.high))))
  ylim_rich <- row_limits(list(rich_depth, rich_temp), "Richness")
  ylim_maxn <- row_limits(list(maxn_depth, maxn_temp), "Total_MaxN")
  xlim_depth <- range(deployments$depth)
  xlim_temp  <- range(deployments$temperature) + c(-0.1, 0.1)

  pA <- prediction_panel(rich_depth, "depth", "Richness", by_var, cols) +
    coord_cartesian(xlim = xlim_depth, ylim = ylim_rich) +
    labs(x = NULL, y = "Predicted species richness") + hide_axis_text("x")
  pB <- prediction_panel(rich_temp, "temperature", "Richness", by_var, cols, jitter = TRUE) +
    coord_cartesian(xlim = xlim_temp, ylim = ylim_rich) +
    labs(x = NULL, y = NULL) + hide_axis_text("x") + hide_axis_text("y")
  pC <- difference_panel(rich_diff, cols) +
    labs(x = NULL, y = "Difference in species richness") + hide_axis_text("x")
  pD <- prediction_panel(maxn_depth, "depth", "Total_MaxN", by_var, cols) +
    coord_cartesian(xlim = xlim_depth, ylim = ylim_maxn) +
    labs(x = "Depth (m)", y = "Predicted total MaxN")
  pE <- prediction_panel(maxn_temp, "temperature", "Total_MaxN", by_var, cols, jitter = TRUE) +
    coord_cartesian(xlim = xlim_temp, ylim = ylim_maxn) +
    labs(x = "Near-bottom temperature (°C)", y = NULL) + hide_axis_text("y")
  pF <- difference_panel(maxn_diff, cols) +
    labs(x = diff_xlab, y = "Difference in total MaxN")

  (pA + pB + pC + pD + pE + pF + plot_layout(ncol = 3)) |
    prediction_legend(levels(deployments[[by_var]]), cols, legend_title)
}

finish_gam_figure <- function(fig) {
  fig + plot_layout(widths = c(1, 0.16)) +
    plot_annotation(tag_levels = list(c(LETTERS[1:6], ""))) &
    tag_inside(0.01, 0.99)
}
```

# 5. Figure 7 – GAMs by location (In / Out)

```{r fig7, fig.width=16, fig.height=10}
figure7 <- finish_gam_figure(gam_figure(
  rich_depth_location, rich_temp_location, diff_rich_location,
  maxn_depth_location, maxn_temp_location, diff_maxn_location,
  by_var = "Location", cols = location_cols, legend_title = "Location:",
  diff_xlab = "Location"))
figure7

save_figure(figure7, "Figure 7", width = 16, height = 10)
```

# 6. Figure 8 – GAMs by canyon

```{r fig8, fig.width=16, fig.height=10}
figure8 <- finish_gam_figure(gam_figure(
  rich_depth_canyon, rich_temp_canyon, diff_rich_canyon,
  maxn_depth_canyon, maxn_temp_canyon, diff_maxn_canyon,
  by_var = "Canyon", cols = canyon_cols, legend_title = "Canyon:",
  diff_xlab = "Canyon"))
figure8

save_figure(figure8, "Figure 8", width = 16, height = 10)
```

# 7. Figure S4 – substrate type by depth

Frequency of occurrence = proportion of the deployments of a depth range in
which the substrate type was recorded; n = number of deployments.

```{r figS4, fig.width=13, fig.height=6}
# Depth bins of 20 m; a depth equal to a bin limit belongs to the lower bin
bin_breaks <- seq(floor(min(deployments$depth) / substrate_bin) * substrate_bin,
                  ceiling(max(deployments$depth) / substrate_bin) * substrate_bin,
                  by = substrate_bin)

substrate <- deployments %>%
  filter(!is.na(Substrate), !is.na(depth)) %>%
  mutate(bin_index = as.integer(cut(depth, breaks = bin_breaks, include.lowest = TRUE)),
         bin_low   = bin_breaks[bin_index],
         depth_bin = paste0(bin_low, "–", bin_low + substrate_bin))

substrate_n <- substrate %>% count(Canyon, bin_low, depth_bin, name = "n_deployments")

substrate_summary <- substrate %>%
  count(Canyon, bin_low, depth_bin, Substrate, name = "n_occurrence") %>%
  left_join(substrate_n, by = c("Canyon", "bin_low", "depth_bin")) %>%
  mutate(occurrence = n_occurrence / n_deployments)

bin_order <- substrate_n %>% distinct(bin_low, depth_bin) %>% arrange(bin_low) %>% pull(depth_bin)
substrate_summary <- substrate_summary %>% mutate(depth_bin = factor(depth_bin, levels = bin_order))
substrate_n       <- substrate_n       %>% mutate(depth_bin = factor(depth_bin, levels = bin_order))

# Substrate types on rows 1-3; the number of deployments of each depth range
# is written in an extra bottom row labelled "n ="
substrate_levels <- levels(substrate_summary$Substrate)
n_row <- 0.25

figureS4 <- ggplot(substrate_summary,
                   aes(depth_bin, as.integer(Substrate))) +
  geom_hline(yintercept = seq_along(substrate_levels), colour = "grey92", linewidth = 0.5) +
  geom_hline(yintercept = 0.6, colour = "grey60", linewidth = 0.4) +
  geom_point(aes(size = occurrence), colour = "grey20", alpha = 0.85) +
  geom_text(data = substrate_n, aes(depth_bin, y = n_row, label = n_deployments),
            size = fig_sizes$annotation + 0.5, inherit.aes = FALSE) +
  facet_grid(~ Canyon, scales = "free_x", space = "free_x") +
  scale_size_continuous(name = "Frequency of\noccurrence:", limits = c(0, 1),
                        breaks = c(0.25, 0.5, 0.75, 1), range = c(2, 9),
                        labels = scales::percent_format()) +
  scale_y_continuous(breaks = c(n_row, seq_along(substrate_levels)),
                     labels = c("n =", substrate_levels),
                     limits = c(n_row - 0.3, length(substrate_levels) + 0.5),
                     expand = expansion(0)) +
  labs(x = "Depth range (m)", y = "Substrate type") +
  theme_bw() + theme_fig +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.background = element_rect(fill = "white", colour = "black", linewidth = 0.6),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.key.size = unit(fig_sizes$legend_key, "lines"))
figureS4

save_figure(figureS4, "Figure S4", width = 13, height = 6)
```

# 8. Figure S5 – redundancy analysis (RDA)

Community: presence/absence of species, Hellinger-transformed. Explanatory
variables: near-bottom temperature, depth and substrate type. Deployments
without fish or with a missing variable are excluded.

```{r rda}
rda_rows <- complete.cases(deployments[, c("depth", "temperature", "Substrate")]) &
            rowSums(community) > 0
rda_data <- deployments[rda_rows, ]
community_hellinger <- decostand(community[rda_rows, ], method = "hellinger")

rda_model <- rda(community_hellinger ~ temperature + depth + Substrate, data = rda_data)
rda_model
set.seed(3)
rda_test_overall <- anova(rda_model, permutations = 999)
rda_test_terms   <- anova(rda_model, by = "margin", permutations = 999)
rda_test_overall
rda_test_terms
vif.cca(rda_model)

write_csv(bind_rows(
  as_tibble(as.data.frame(rda_test_overall), rownames = "Term") %>% mutate(Test = "overall"),
  as_tibble(as.data.frame(rda_test_terms),   rownames = "Term") %>% mutate(Test = "by margin")),
  file.path(fig_dir, "Table_RDA_tests.csv"))
```

```{r figS5, fig.width=10, fig.height=9}
site_scores <- as_tibble(scores(rda_model, display = "sites")) %>% bind_cols(rda_data)

arrows_cont <- as.data.frame(scores(rda_model, display = "bp"))
arrows_cont <- arrows_cont[rownames(arrows_cont) %in% c("depth", "temperature"), ]
arrows_sub  <- as.data.frame(scores(rda_model, display = "cn"))
env_arrows  <- bind_rows(arrows_cont, arrows_sub) %>%
  rownames_to_column("Variable") %>%
  mutate(Label = recode(Variable, depth = "Depth", temperature = "Temperature"),
         Label = str_remove(Label, "^Substrate"),
         RDA1  = RDA1 * rda_arrow_scale,
         RDA2  = RDA2 * rda_arrow_scale)

variance <- summary(rda_model)$cont$importance["Proportion Explained", c("RDA1", "RDA2")]
axis_title <- function(axis) sprintf("%s (%.1f%%)", axis, 100 * variance[[axis]])

location_lines <- c("In" = "solid", "Out" = "dashed")

figureS5_plot <- ggplot(site_scores, aes(RDA1, RDA2)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  stat_ellipse(aes(group = Location, linetype = Location), data = filter(site_scores, Location == "In"),
               colour = location_cols[["In"]], level = 0.95, linewidth = 0.8) +
  stat_ellipse(aes(group = Location, linetype = Location), data = filter(site_scores, Location == "Out"),
               colour = location_cols[["Out"]], level = 0.95, linewidth = 0.8) +
  geom_point(aes(colour = Canyon, shape = Location), size = 3.5, alpha = 0.85) +
  geom_segment(data = env_arrows, aes(x = 0, y = 0, xend = RDA1, yend = RDA2),
               arrow = arrow(length = unit(0.22, "cm")), linewidth = 0.6, colour = "black") +
  geom_text_repel(data = env_arrows, aes(RDA1, RDA2, label = Label), family = font_family,
                  fontface = "bold", size = fig_sizes$annotation + 0.5, point.padding = 0.3,
                  segment.colour = "grey60", segment.size = 0.3, max.overlaps = Inf, seed = 4) +
  scale_colour_manual(values = canyon_cols, guide = "none") +
  scale_shape_manual(values = c("In" = 16, "Out" = 17), guide = "none") +
  scale_linetype_manual(values = location_lines, guide = "none") +
  scale_x_continuous(labels = label_2dec) +
  scale_y_continuous(labels = label_2dec) +
  labs(x = axis_title("RDA1"), y = axis_title("RDA2")) +
  theme_panel

legend_rda <- legend_panel(
  ggplot(tibble(Canyon   = factor(canyon_levels, levels = canyon_levels),
                Location = factor(c("In", "Out", "Out"), levels = location_levels),
                x = 0, y = 0), aes(x, y)) +
    geom_point(aes(colour = Canyon), alpha = 0, size = 3.5) +
    geom_point(aes(shape = Location), alpha = 0, size = 3.5) +
    geom_line(aes(linetype = Location, group = Location), alpha = 0) +
    scale_colour_manual(values = canyon_cols, name = "Canyon:",
                        guide = guide_legend(order = 1, override.aes = list(alpha = 1, shape = 16))) +
    scale_shape_manual(values = c("In" = 16, "Out" = 17), name = "Location:",
                       guide = guide_legend(order = 2, override.aes = list(alpha = 1))) +
    scale_linetype_manual(values = location_lines, name = "Ellipse (95%):",
                          guide = guide_legend(order = 3, override.aes = list(
                            alpha = 1, linewidth = 0.8, colour = unname(location_cols)))),
  scale = 1.1)

figureS5 <- figureS5_plot + legend_rda + plot_layout(widths = c(1, 0.25))
figureS5

save_figure(figureS5, "Figure S5", width = 10.5, height = 9)
```

# Save the workspace and the R session information

```{r save-session}
save.image(file = file.path(out_dir, "03_Bio_Environmental.RData"))
sessionInfo()
```
